Basic Model formulation

### 2015-04-14 documentation/model_formulation.Rmd
update model formulation: bifurcation section

defparms <- list(
    m = 0.05,
    r = 1,
    b = 0.9, 
    K = 0.9, 
    a = 0.6,
    h = 50,
    L = 6, 
    alpha = 0.3,
    q = 0.5,
    c = 0,
    f = 0,
    v = 0,
    p = 0

@@

If those effects coincide, potentially even with the global feedbacks, the intersections are difficult to estimate.

@@

parms <- list(
    m = 0.05,  # intrinsic mortality of plants 
    r = 1,   # growth rate
    f = 0.8, #  local facilitation
    b = 0.4,  # environmental quality inverse of aridity
    K = 0.9,  # carrying capacity
    alpha = 0.1, # water runoff
    c = 0.2, # local competition
    a = 0.6, # search efficiency
    v = 0.1, # attraction effect
    h = 50, # handling time (one individual on landscape unit)
    p = 0.8, # protection effect
    L = 6, # livestock units per landscape unit 
    q = 1 # search ineffiency at low cover
  )

@@

parms <- list(
    m = 0.05,  # intrinsic mortality of plants 
    r = 1,   # growth rate
    f = 0.8, #  local facilitation
    b = 0.3,  # environmental quality inverse of aridity
    K = 0.9,  # carrying capacity
    alpha = 0, # water runoff
    c = 0.2, # local competition
    a = 0.6, # search efficiency
    v = 0.1, # attraction effect
    h = 50, # handling time (one individual on landscape unit)
    p = 0.8, # protection effect
    L = 6, # livestock units per landscape unit 
    q = 1 # search ineffiency at low cover
  )

attractor(parms)

@@

The following shows bifurcation diagrams over an environmental gradient, $b$, as well as over a gradient of livestock density, $L$.

```{r, echo=FALSE, fig.height = 3}
par(mfrow = c(1,2))
bifurcation(parms, "L", c(0,12))

bifurcation(parms,"b", c(0,1))

>>>>

defparms <- list(  
    m = 0.05,
    r = 1,
    b = 0.9, 
    K = 0.9, 
    a = 0.6,
    h = 100,
    L = 16, 
    alpha = 0.1,
    q = 1,
    c = 0,
    f = 0,
    v = 0,
    p = 0

@@

If those effects coincide, potentially even with the global feedbacks, the intersections are difficult to estimate visually.

@@

parms <- list(
    m = 0.05,  # intrinsic mortality of plants 
    r = 1,   # growth rate
    f = 0.8, #  local facilitation
    b = 0.4,  # environmental quality inverse of aridity
    K = 0.9,  # carrying capacity
    alpha = 0, # water runoff
    c = 0.2, # local competition
    a = 0.2, # search efficiency
    v = 0.8, # attraction effect
    h = 50, # handling time (one individual on landscape unit)
    p = 0.99, # protection effect
    L = 5, # livestock units per landscape unit 
    q = 0 # search ineffiency at low cover
  )

@@

The following shows bifurcation diagrams of the mean field approximation over an environmental gradient, $b$, as well as over a gradient of livestock density, $L$.

```{r, echo=FALSE, fig.height = 3 }
par(mfrow = c(1,2))
parms$q = 0.4
bifurcation(parms, "L", c(0,12)) -> over_L

bifurcation(parms,"b", c(0,1)) -> over_b

@@

The same graph using the pair-approximation.

```{r, echo=FALSE, fig.height = 3}
par(mfrow = c(1,2))
bifurcation(parms, "L", pairapprox = TRUE, meanfield = FALSE, c(0,12)) -> over_L

bifurcation(parms,"b", pairapprox = TRUE, meanfield = FALSE, c(0,1)) -> over_b

The bifurcation diagram visualises the attractor only along the total cover which is only an incomplete visualisation, since the steady state is described by both, global vegetation cover $\rho_1$ and average local cover $q_{1|1}$ (or alternatively the cover of vegetated pairs $\rho_{11}$ but this is rather unintuitive). To visualise this, a three dimensional graph is necessary.

```{r, echo=FALSE, fig.height = 5}

persp(0:1,0:1, matrix(c(0,0,0,0), ncol = 2), zlim = c(0,1), ticktype = "detailed", zlab = "vegetation cover", ylab = "local cover" , xlab = "b", theta = -65, phi = 19, r = sqrt(3), d = 1) -> p

  points(trans3d(over_b$pairapprox$b, over_b$pairapprox$rho_1, rep(0, length(over_b$pairapprox$rho_1)), pmat = p), col = "#D0D0D0", pch = 20,  cex = 0.5)
  points(trans3d(over_b$pairapprox$b,rep(1, length(over_b$pairapprox$rho_1)), over_b$pairapprox$rho_11/over_b$pairapprox$rho_1,  pmat = p), col = "#D0D0D0", pch = 20,  cex = 0.5)  
  points(trans3d(rep(1, length(over_b$pairapprox$rho_1)),over_b$pairapprox$rho_1, over_b$pairapprox$rho_11/over_b$pairapprox$rho_1,  pmat = p), col = "#D0D0D0", pch = 20,  cex = 0.5)

  points(trans3d(over_b$pairapprox$b, over_b$pairapprox$rho_1, over_b$pairapprox$rho_11/over_b$pairapprox$rho_1, pmat = p), pch = 20,  cex = 0.5)

Decomposing it into its dimensions is easier to print.

```{r, echo = FALSE}
over_b$pairapprox$q_11 <- over_b$pairapprox$rho_11/over_b$pairapprox$rho_1
over_b$pairapprox$q_11[is.na(over_b$q_11)] <- 0

par(mfrow =c(1,2))
  plot(over_b$pairapprox$b, over_b$pairapprox$rho_1, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
  plot(over_b$pairapprox$b, over_b$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
```

The direct comparison of global cover vs. local cover indicates the amount of clustering along the stress gradient and at the critical thresholds. 
We are going to compare these plots for a range of models to see how the spatial configuration changes with increasing stress under different mechanism scenarios.

```{r, echo=FALSE, fig.width = 3}
par(mfrow = c(1,1))
  plot(over_b$pairapprox$rho_11, over_b$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
abline(a = 0, b = 1, lty = 3, col = "#D0D0D0")
points(0,0, pch=20, col = "white", cex = 2)
 points(0,0, pch=20, cex = 0.5)
